---
title: "Linear algebra"
author: "Neko_Chan666"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    number_sections: TRUE
    theme: journal
    df_print: paged
    css: style.css
---

# Linear algebra
It is often convenient to think biological datasets as 'matrices', as values are of same set of variables (e.g., levels of expression of different genes) in a set of biological samples (e.g., cell lines with different treatment protocols). So performing mathematical operations on matrices (termed as 'linear algebra') is worth learning.
A matrix is a grid of numbers, each of which can be referred to by row and column indices.
Let's define a matrix A of dimension 3x2. (it has 3 rows and 2 columns)

$$\begin{bmatrix}
A_{11}&A_{12} \\
A_{21}&A_{22} \\
\end{bmatrix}$$

## Matrix multiplication
If we multiply two matrices, then a new matrix with the number of row of the first matrix and the number of column of the second matrix will be formed.

$$ If\ AB = C,\ C_{ij} = \sum_{k = 1}^{p} A_{ik}\ *\ B_{kj}$$

If A is an (n x p) matrix and B is a (p x m) matrix, then AB will be an (n x m) matrix.

Let's make a matrix in R.

## Matrix in R
Matrix A with dimensions (3 x 2).
```{r}
A <- matrix(c(1, 2, 3, 4, 5, 6), nrow = 3, ncol = 2)

A
```

Let's make another matrix B with dimensions (2 x 3).
```{r}
B <- matrix(c(1, 2, 3, 4, 5, 6), nrow = 2, ncol = 3)

B
```

## Matrix multiplication in R
To multiply these two matrices we need an operator "%*%".
```{r}
C <- A%*%B

C
```

As you can see the resultant matrix C is a (3 x 3) matrix.

# Why?
As we already know, a set of values sampled from a population of values modeled as a 'random variable'. And if multiple measurements are obtained from the same objects, the set of vectors representing each measurement's values can be treated as a matrix.

> A matrix with just one row or one column is usually referred to as a vector.

This is usually the case for the datasets we will be working with ^_^