---
title: "R Basics One"
author: "Neko_Chan666"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    number_sections: TRUE
    theme: journal
    df_print: paged
    css: style.css
---

# R workspace
As you know R is a statistical programming environment that can store objects in it's workspace. If we create any object in current workspace, then it can be viewed by
```{r}
ls()
```
As you can see there are no objects at first.
So lets create an object.

## Creating object
```{r}
x <- 1

ls()
```
Now there is an object called 'x' with a value 1. Lets try to remove it.

## Removing object
```{r}
rm(x)

ls()
```
After using rm(<obj_name>) function, there is no object.

## Saving workspace
To save your workspace into a file use
```{r}
save.image("workspace.Rdata")
```
This will save our current workspace ^_^

# Data handling
The objects that can be manipulated in a wide variety of ways in R. So lets start @_#

## Basic data types
    1. Character ~> These are the combination of alpha-numeric character that are quoted by "".
    2. Numeric ~> A number value.
    3. Logical ~> TRUE or FALSE

## Data structures

### Vectors
In the context of R, a vector is a simple one-dimensional list of objects which are all of the same basic type.
Now to create a vector we have to use 'concatenate' function c().
```{r}
vec1 <- c(1, 2, 3) # All integer

vec1

vec2 <- c(1, 3, 4, 5.4) # If vector contains a float

vec2

vec3 <- c(2:10) # Vector of sequential integers

vec3
```

it's a bit confusing but the indexing of vectors starts with 1. So to retrieve an object from the vector we will use the indexing.
```{r paged.print=TRUE}
vec1[1]

vec2[vec1]
```

The second retrieval can be a little confusing so let me tell. 'vec1' contains 3 objects (1, 2, 3) so when we put 'vec1' inside 'vec2' then (1, 2, 3) become indexes of (1, 3, 4, 5.4). After running 'vec2[vec1]' we get (1, 3, 4) as outputs.

> What if you use negative indexing?

> Well that's a valid question. Negative indices don't return the indexed element(s).

```{r paged.print=TRUE}
vec2[-1] # index 1 ~> 1; there is no 1

vec2[-3] # index 3 ~> 4; and there is no 4
```

### Arrays
Arrays are multi-dimensional vectors, meaning it's like a table of objects of same basic data type.
Arrays can be create in R using the function array() that is filled with values in order of the dimensions (like row first, column second...)
```{r}
arr <- array(c(3:17), dim = c(3, 5))

arr
```

Remember, if the number of elements exceed the dimension then after filling the dimension elements will be lost o_O
Just like vectors, arrays are also indexed with square brackets with row first and column second.
```{r paged.print=TRUE}
arr[1, 3] # first row and third column

arr[2,] # second row

arr[, 5] # fifth column

arr[,c(1,3)] # first and third column
```

### Lists
List and vector both data structures can store elements but in a vector all basic data types should have to be same whereas in a list there can exist different data types. Lets use it e_e
```{r}
myList <- list(1, 2, 3)

myList
```

Now to index one of the elements from within a list, we have to [[index]].
```{r}
myList[[1]]
```

Lets change the index [[1]] with a collection.
```{r}
myList[[1]] <- c(32, 43, 9)

myList
```

> As you can see, an individual element of a list in R may be a list itself. !_!

For this property, list is also known as a 'recursive' data structure.
```{r}
myList[[2]] <- list("gene A", "gene B", "gene C")

myList
```

We can also retrieve an element by names.
Lets say we name an element 'neko'.
```{r paged.print=TRUE}
myList[[3]] <- list(21, 33, neko = 6)

myList

myList$neko # Remember, we put the list into the index [[3]] not overwrite the whole list

myList[[3]]$neko
```

### Data frames
Yes you guessed right!!!
Data frames also belong to data structures.
A data frame is a bit of a cross between an array and a list.

> Data frame is essentially a list in which the elements must all be the vector of the same length, although they can be of different basic data types.

Basically it's a table of different data types that may be considered related. 
Lets create a data frame with 2 columns, one of which represents the numbers from 1 to 9 and other in their textual form.
```{r}
myFrame <- data.frame(no. = c(1:9), names = c("one", "two", "three", "four", "five", "six", "seven", "eight", "nine"))

myFrame
```

This is the standard format in R for tables of data, as it can represent multiple attributes of a set of things.

## Data input/output
Now after making a data frame 'myFrame', lets save the object from the workspace to a file (lets say, "myFrame.txt").

OOh... before that, lets create a directory first @_@!

### Directory creation
```{r}
directory <- "./R_basic_one"

if (dir.exists(directory)){
  cat("The directory already exists...")
} else {
  dir.create(directory, showWarnings = TRUE, recursive = FALSE, mode = "0777")
  cat("Created !!!")
}
```
As you can see, here we use two functions "dir.create" and "dir.exists". The function dir.create() function creates a new directory with a specified path whereas dir.exists() function checks whether the directory exists or not.

> dir.create(path, showWarnings = TRUE, recursive = FALSE, mode = "0777")

#### Parameters
    1. path ~> Relative or absolute path of the directory, the directory we want to create.
    2. showWarnings ~> It is a logical argument means if there is a failure then it will print FALSE.
    3. recursive ~> It is a logical argument that is very similar to "mkdir -p". It means that "mkdir" can only create a folder in the current working directory but if we want to create a directory in another location then the we have to use "mkdir -p".
    4. mode ~> Giving full permission of 'rwx' for 'user, group and others'. Means giving full permission.
    
#### Return Value
The dir.create() function returns an invisibly logical vector indicating if the operation succeeded for each of the files attempted.

### Storing as a file
Now lets write a file named "myFrame.txt" containing our data frame.

```{r}
myFile <- "./R_basic_one/myFrame.txt"

write.table(myFrame, file = myFile, sep = "\t", quote = FALSE, row.names = FALSE)
```

> write.table(dataframe, file = "file_name.txt", sep = "\t", quote = FALSE, row.name = FALSE)

Here we replace "./R_basic_one/myFrame.txt" with "file_name.txt" ?_?

#### Parameters
    1. sep ~> This will separate our table with some input string. Like here we use "\t", it means a tab, which is useful for Excel.
    2. quote ~> By default R places double quotes " around all characters in the table. So we mark quote as FALSE. It's a good practice to remove these.
    3. row.names / col.names ~> By default write.table() funtion writes an additional row and column names in the file. So to get rid of these, we specify row.names = FALSE.
    
### Reading a file
```{r}
newFrame <- read.table(file = myFile, sep = "\t", header = TRUE)

newFrame
```

#### Parameters
    1. file ~> This indicates the date frame we store in our secondary memory.
    2. sep ~> Stands for separator.
    3. header ~> It says that the first row of the file should be used as the column headers.
  
```{r}
newFrame$names
```